---
title: "Installation of System Requirements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation of System Requirements}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

Configuring a workstation to run mldash requires installing a few major components: R; Python; Java. Of course, mldash runs tidymodels natively in R. However, mldash will also run Python (e.g. Prophet) & Java (Weka) models so this guide will outline steps on how to build a working environment.

Depending on your operating system, you can use various system tools or package managers to ease the installation process.

At the end of this guide, you should have a workstation meets the following goals:

* Runs R and R Studio
* Has all the required R library dependencies installed
* R pointing to a functioning Python environment with dependencies
* R pointing to a functioning Java run time environment

**Note:** these directions have been tested on macOS 12.6 (Monterey on a M1 MacBook, ARM processor) and Red Hat Enterprise Linux (RHEL) 9. These steps have not been tested on other platforms.


***

## Installing R & RStudio

You can download and install the R binary from the [Crane Project home page](https://cran.r-project.org). Next you'll need to install [RStudio](https://posit.co), as your integrated development environment.

### macOS

In an RStudio session, you can install required libraries in the console window:

```{r m1_dependencies, eval=FALSE}
install.packages("glmnet")
install.packages("brulee")
install.packages("fastDummies")
install.packages("kknn")
install.packages("plsmod")
install.packages("remotes")
install.packages("baguette")
install.packages("libcoin")
install.packages("earth")
install.packages("dbarts")
install.packages("xgboost")
install.packages("forecast")
install.packages("modeltime")
```

#### Installing the mix0mics Library

To install the mix0mics library, you need to install the [BiocManager package](http://www.bioconductor.org/packages/release/bioc/html/mixOmics.html).

```{r mix0mics_dependency, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("mixOmics")
```



### Red Hat Enterprise Linux

Installing R dependencies is very similar:

```{r rhel_dependencies, eval=FALSE}
install.packages(c("baguette","libcoin","earth","dbarts","xgboost","forecast","modeltime","glmnet","brulee","fastDummies","kknn","plsmod"))

install.packages(c("BiocManager"),type="binary")

install.packages(c("poissonreg","pscl","ranger","kernlab","mda","discrim","sda","sparsediscrim","klaR","LiblineaR","naivebayes","rules","MASS","baguette"),type="binary")


install.packages(c("dbarts","discrim","earth","fastDummies","glmnet","keras","kernlab","kknn","klaR","mda","mgcv","mixOmics","naivebayes","nnet","parsnip",
"plsmod","poissonreg","randomForest","ranger","rpart","rstanarm","rules","sda","sparsediscrim","stats","xgboost","xrf"),type="binary")
```

## Installing Python - Method #1 with Miniconda

Many of the models will require Python which is executed using the `reticulate` package. I, personally, have found the installation and configuration of Python to be frustrating, especially on a Mac M1. However, as of this writing, the following works (on my system). First, install these packages from Github to ensure the latest version.

```{r install_pythong_pkgs, eval=FALSE}
remotes::install_github(sprintf("rstudio/%s", c("reticulate", "tensorflow", "keras", "torch")))
```

If you have previously installed Miniconda, it is helpful to start from a clean slate. 

```{r miniconda_uninstall, eval=FALSE}
reticulate::miniconda_uninstall()
```

We can then install Miniconda using the following command:

```{r install_miniconda, eval=FALSE}
reticulate::install_miniconda()
```

Once installed, we can create a conda environment:

```{r conda_create, eval=FALSE}
reticulate::conda_create("mldash")
```

And then make it active (note sure if it is necessary to do this for all three packages, but it doesn't hurt):

```{r use_condaenv, eval=FALSE}
reticulate::use_condaenv("mldash")
tensorflow::use_condaenv("mldash")
keras::use_condaenv("mldash")
```

Although there are utility functions to install `keras`, `tensorflow`, and `torch` from their respective packages, I found them to not always work as expected. The `conda_install` function will ensure the Python packages are installed into the correct environment. Note that as of this writing, `pytorch` still does not have a Mac M1 native version so some predictive models will not work on that platform.

```{r conda_install, eval=FALSE}
reticulate::conda_install("mldash", 
			  c("jupyterlab", "pandas", "statsmodels",
			    "scipy", "scikit-learn", "matplotlib",
			    "seaborn", "numpy", "pytorch", "tensorflow"))
```

Lastly, ensure that `reticulate` uses the correct Python by setting the `RETICULATE_PYTHON` environment variable (this can also be put in your `.Renviron` file to be used across sessions, though I avoid doing that so I can use different Python paths for different projects).

```{r RETICULATE_PYTHON, eval=FALSE}
Sys.setenv("RETICULATE_PYTHON" = "~/miniforge3/envs/mldash/bin/python")
```


## Installing Python - Method #2 with Mambaforge

On both macOS and RHEL, Python 3.x is pre-installed (either as python or python3) but you'll really need a lot of Python modules. You can get the Mambaforge installer from (Mambaforge's Github page)[https://github.com/conda-forge/miniforge#mambaforge].

Now you can use it to install the many model library dependencies.

```{sh mac_mamba, eval=FALSE}
mamba activate mldash
mamba install -y jupyterlab
mamba install -y pandas
mamba install -y statsmodels 
mamba install -y scipy
mamba install -y scikit-learn
mamba install -y matplotlib 
mamba install -y seaborn
mamba install -y numpy
mamba install -y pytorch
mamba install -y tensorflow
```

As in the previous method, set an environment variable for your Mamba-Python binary:

```{r RETICULATE_PYTHON_MAMBA, eval=FALSE}
Sys.setenv("RETICULATE_PYTHON" = "~/mambaforge/envs/mldash/bin/python")
```



## Installing Java (macOS)

You will need to install a Java version 8 development kit compiled for an ARM processor. One that I found to work is from [Azul](https://www.azul.com/downloads/?package=jdk#download-openjdk).

Follow the install directions and explicitly set an environment variable for JAVA_HOME, otherwise R will try to use the system's default pre-installed Java virtual machine.

```{r JAVA, eval=FALSE}
Sys.setenv(JAVA_HOME='/Library/Java/JavaVirtualMachines/zulu-8.jdk/Contents/Home/jre/')
```


## Installing Java (RHEL)

In Linux, you can use the regular yum package manager and install OpenJDK version 8 on the command line.

```{r java_rhel, eval=FALSE}
$ sudo yum install -y java-1.8.0-openjdk
```
Unlike with macOS, you shouldn't need to set a JAVA_HOME since Java is normally part of the system $PATH variable and normally linked to the correct version. Of course, if you have multiple Java versions, then go ahead and specify JAVA_HOME for safety.


## Next Steps

Now that you have R, Python and Java installed can start [Running Predictive Models](https://jbryer.github.io/mldash/#running-predictive-models).