---
title: "Installation of System Requirements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mldash)
```


# Introduction

Configuring a workstation to run mldash requires installing and supporting a few major components: R; Python; Java. Of course, R runs mldash and many R-native models. However, mldash will also run Python & Java models so this guide will give steps on how to building a working environment.

Depending on your particular operating system, you can use various system tools, package managers to ease the installation process.

At the end of this guide, you should have a workstation meets the following goals:
 * Runs R and R Studio
 * Has all the required R library dependencies installed
 * R pointing to a working Python installation
 * R pointing to a working Java 8 installation

# Installing R & RStudio

Aside from R and RStudio, you'll need to install many model dependencies.

### M1 Mac

### Red Hat Linux
[ install.packages(c("baguette","libcoin","earth","dbarts","xgboost","forecast","modeltime","glmnet","brulee","fastDummies","kknn","plsmod"))

install.packages(c("BiocManager"),type="binary")

install.packages(c("poissonreg","pscl","ranger","kernlab","mda","discrim","sda","sparsediscrim","klaR","LiblineaR","naivebayes","rules","MASS","baguette"),type="binary")

,"dbarts","discrim","earth","fastDummies","glmnet","keras","kernlab","kknn","klaR","mda","mgcv","mixOmics","naivebayes","nnet","parsnip",
"plsmod","poissonreg","randomForest","ranger","rpart","rstanarm","rules","sda","sparsediscrim","stats","xgboost","xrf"),type="binary")
]


### Windows


# Installing Python


## System level installation

### Windows

```{bash, eval=FALSE}
```

```{python, eval=FALSE}
```

### Mac

[M1 - Install Mambaforge from https://mamba.readthedocs.io/]


### Ubuntu

### Red Hat Linux
[Install from https://conda-forge.org]

## Using Reticulate

### Python

Many of the models will require Python which is executed using the `reticulate` package. I, personally, have found the installation and configuration of Python to be frustrating, especially on a Mac M1. However, as of this writing, the following works (on my system). First, install these packages from Github to ensure the latest version.

```{r install_pythong_pkgs, eval=FALSE}
remotes::install_github(sprintf("rstudio/%s", c("reticulate", "tensorflow", "keras", "torch")))
```

If you have previously installed Miniconda, it is helpful to start from a clean slate. 

```{r miniconda_uninstall, eval=FALSE}
reticulate::miniconda_uninstall()
```

We can then install Miniconda using the following command:

```{r install_miniconda, eval=FALSE}
reticulate::install_miniconda()
```

Once installed, we can create a conda environment:

```{r conda_create, eval=FALSE}
reticulate::conda_create("mldash")
```

And then make it active (note sure if it is necessary to do this for all three packages, but it doesn't hurt):

```{r use_condaenv, eval=TRUE}
reticulate::use_condaenv("mldash")
tensorflow::use_condaenv("mldash")
keras::use_condaenv("mldash")
```

Although there are utility functions to install `keras`, `tensorflow`, and `torch` from their respective packages, I found them to not always work as expected. The `conda_install` function will ensure the Python packages are installed into the correct environment. Note that as of this writing, `pytorch` still does not have a Mac M1 native version so some predictive models will not work on that platform.

```{r conda_install, eval=FALSE}
reticulate::conda_install("mldash", 
						  c("jupyterlab", "pandas", "statsmodels",
						    "scipy", "scikit-learn", "matplotlib",
						    "seaborn", "numpy", "pytorch", "tensorflow"))
```

Lastly, ensure that `reticulate` uses the correct Python by setting the `RETICULATE_PYTHON` environment variable (this can also be put in your `.Renviron` file to be used across sessions, though I avoid doing that so I can use different Python paths for different projects).

```{r RETICULATE_PYTHON, eval=FALSE}
Sys.setenv("RETICULATE_PYTHON" = "~/miniforge3/envs/mldash/bin/python")
```




# Installing Java

### M1 Mac
[install Azul j8 jdk]

### Red Hat Linux
[sudo yum install -y java-1.8.0-openjdk]



