---
title: "Installation of System Requirements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mldash)
```


# Introduction

Configuring a workstation to run mldash requires installing and supporting a few major components: R; Python; Java. Of course, R runs mldash and many R-native models. However, mldash will also run Python & Java models so this guide will give steps on how to building a working environment.

Depending on your particular operating system, you can use various system tools, package managers to ease the installation process.

At the end of this guide, you should have a workstation meets the following goals:
 * Runs R and R Studio
 * Has all the required R library dependencies installed
 * R pointing to a working Python installation
 * R pointing to a working Java 8 installation

# Installing R & RStudio

Aside from R and RStudio, you'll need to install many model dependencies.

### M1 Mac

### Red Hat Linux
[ install.packages(c("baguette","libcoin","earth","dbarts","xgboost","forecast","modeltime","glmnet","brulee","fastDummies","kknn","plsmod"))

install.packages(c("BiocManager"),type="binary")

install.packages(c("poissonreg","pscl","ranger","kernlab","mda","discrim","sda","sparsediscrim","klaR","LiblineaR","naivebayes","rules","MASS","baguette"),type="binary")

,"dbarts","discrim","earth","fastDummies","glmnet","keras","kernlab","kknn","klaR","mda","mgcv","mixOmics","naivebayes","nnet","parsnip",
"plsmod","poissonreg","randomForest","ranger","rpart","rstanarm","rules","sda","sparsediscrim","stats","xgboost","xrf"),type="binary")
]


### Windows


# Installing Python

## System level installation

### Windows

```{bash, eval=FALSE}
```

```{python, eval=FALSE}
```

### Mac

[M1 - Install Mambaforge from https://mamba.readthedocs.io/]


### Ubuntu

### Red Hat Linux
[Install from https://conda-forge.org]

## Using Reticulate


# Installing Java

### M1 Mac
[install Azul j8 jdk]

### Red Hat Linux
[sudo yum install -y java-1.8.0-openjdk]



